Resonance Fourier Transform:\\ Unitary Lens for Compression, Crypto,\\ and Quantum-Inspired Computing

### 2025-12-15 docs/proofs/PHI_RFT_PROOFS.tex
Update Canonical RFT definition to Gram-normalized basis and align docs/proofs

\subsection{Canonical QR-Based RFT}

The \emph{canonical} RFT is constructed via QR decomposition of a golden-ratio weighted kernel.

\begin{definition}[Golden Resonance Kernel]\label{def:kernel}
The golden resonance kernel $K \in \R^{n \times n}$ is defined by:
\begin{equation}
K_{ij} = \phi^{|i-j|} \cdot \cos\left(\frac{\phi \cdot i \cdot j}{n}\right), \quad i,j \in \{0,\ldots,n-1\}
\end{equation}
where $\phi = (1+\sqrt{5})/2$ is the golden ratio.
\end{definition}

\begin{definition}[Canonical RFT Matrix]\label{def:canonical}
The canonical RFT matrix $U \in \C^{n \times n}$ is the $Q$ factor from QR decomposition:
\begin{equation}
K = UR \quad \text{where } U\dg U = I_n
\end{equation}
\end{definition}

\begin{theorem}[Canonical RFT Unitarity]\label{thm:canonical_unitary}
The canonical RFT matrix $U$ is exactly unitary.
\end{theorem}

\begin{proof}
By definition of QR decomposition, the $Q$ factor has orthonormal columns.
\end{proof}

>>>>

\subsection{Canonical Gram-Normalized RFT}

The \emph{canonical} RFT is constructed via Gram-matrix normalization (symmetric orthogonalization) of an irrational-frequency exponential basis.

\begin{definition}[Irrational Frequency Basis]\label{def:kernel}
The raw basis matrix $\Phi \in \C^{n \times n}$ is defined by:
\begin{equation}
\Phi_{tk} = \frac{1}{\sqrt{n}} \exp\left(j 2\pi f_k t\right), \quad t,k \in \{0,\ldots,n-1\}
\end{equation}
where $f_k = \text{frac}((k+1)\phi)$ are golden-ratio frequencies folded to $[0,1)$.
\end{definition}

\begin{definition}[Canonical RFT Matrix]\label{def:canonical}
The canonical RFT matrix $\widetilde{\Phi} \in \C^{n \times n}$ is obtained via Gram-matrix normalization:
\begin{equation}
\widetilde{\Phi} = \Phi (\Phi\dg \Phi)^{-1/2}
\end{equation}
where $(\Phi\dg \Phi)^{-1/2}$ is the inverse square root of the Gram matrix $G = \Phi\dg \Phi$.
\end{definition}

\begin{theorem}[Canonical RFT Unitarity]\label{thm:canonical_unitary}
The canonical RFT matrix $\widetilde{\Phi}$ is exactly unitary.
\end{theorem}

\begin{proof}
Let $G = \Phi\dg \Phi$ be the Gram matrix. Since $\Phi$ is full-rank (generically true), $G$ is Hermitian positive definite.
\begin{equation}
\widetilde{\Phi}\dg \widetilde{\Phi} = (G^{-1/2})\dg \Phi\dg \Phi G^{-1/2} = G^{-1/2} G G^{-1/2} = I_n
\end{equation}
This construction (Loewdin orthogonalization) yields the unique unitary matrix closest to $\Phi$ in the Frobenius norm.
\end{proof}

### 2025-12-15 docs/research/THE_PHI_RFT_FRAMEWORK_PAPER.tex
Update Canonical RFT definition to Gram-normalized basis and align docs/proofs

\title{The $\Phi$-RFT Framework: Seven Unitary Transforms for Hybrid Discrete-Continuous Signal Processing}
\author{Luis M. Minier \\ \textit{Independent Researcher} \\ \href{mailto:luisminier79@gmail.com}{luisminier79@gmail.com}}
\date{November 24, 2025}

@@

\begin{abstract}
We present the $\Phi$-Resonant Fourier Transform ($\Phi$-RFT) framework, a family of seven rigorously unitary transforms derived from Golden Ratio phase modulation. The framework exactly diagonalizes systems governed by irrational resonances and, via Theorem~10, resolves the long-standing ``ASCII Bottleneck'' that prevented continuous spectral transforms from efficiently representing symbolic data. This manuscript consolidates the mathematical proofs, implementation architecture, verification test suite, and MATLAB-oriented figure generation pipeline that together establish enablement for the QuantoniumOS wave-computing stack.
\end{abstract}

@@

>>>>

	itle{The Quantonium Resonance Fourier Transform: Deterministic Irrational-Frequency Frame Normalization and Empirical Validation}
\author{Luis M. Minier \\ \textit{Independent Researcher} \\ \href{mailto:luisminier79@gmail.com}{luisminier79@gmail.com}}
\date{December 15, 2025}

@@

\begin{abstract}
We present the Quantonium Resonance Fourier Transform (RFT) as a reproducible, unitary transform stack centered on deterministic Golden Ratio ($\phi$) structure. A key contribution is a mathematically complete treatment of an irrational-frequency exponential construction at finite transform size $N$: because irrationally spaced exponentials are generally non-orthogonal for finite $N$, correlation inversion is not valid without correction. We therefore introduce Gram-matrix normalization, $\widetilde{\Phi}=\Phi(\Phi^H\Phi)^{-1/2}$, which yields a unitary (Parseval/tight) basis at finite $N$, and we provide dual-frame coefficient extraction as an exact alternative.

We further report empirical validation artifacts from the QuantoniumOS repository: (i) machine-precision unitarity tests, (ii) finite-$N$ reconstruction benchmarks for frame-corrected and Gram-normalized $\phi$-grid bases, and (iii) large-$N$ (up to $N=4096$) coherence and real-data coefficient diagnostics. Separately, we retain a closed-form $O(N\log N)$ $\phi$-phase FFT operator that is hardware-friendly; the paper distinguishes this fast unitary operator from the irrational-frequency frame-normalized construction.
\end{abstract}

@@

\subsection{Deterministic Irrational-Frequency Frame Normalization}

Define a square exponential basis $\Phi\in\mathbb{C}^{N\times N}$ using deterministic irrational frequencies (folded into the discrete-time fundamental band). For example, with $f_k=\operatorname{frac}((k+1)\phi)$ and $n\in\{0,\dots,N-1\}$,
\[
\Phi[n,k]=\frac{1}{\sqrt{N}}e^{i2\pi f_k n}.
\]
At finite $N$, such irrational-frequency exponentials are generally not exactly orthogonal, i.e. $\Phi^H\Phi\neq I$.

Let $G=\Phi^H\Phi$ denote the Gram matrix. If $\Phi$ is full rank, then $G$ is Hermitian positive definite and $G^{-1/2}$ exists.

\begin{theorem}[Gram-normalized unitary basis]
If $\Phi$ is full rank and $G=\Phi^H\Phi$, then $\widetilde{\Phi}=\Phi G^{-1/2}$ satisfies $\widetilde{\Phi}^H\widetilde{\Phi}=I$.
\end{theorem}

\begin{proof}
\[
\widetilde{\Phi}^H\widetilde{\Phi}=(G^{-1/2})^H\Phi^H\Phi G^{-1/2}=G^{-1/2}GG^{-1/2}=I.
\]
\end{proof}

	extbf{Dual-frame coefficients.} If one keeps the raw basis $\Phi$, exact synthesis coefficients for $x=\Phi X$ are given by the square dual-frame solve
\[
X=(\Phi^H\Phi)^{-1}\Phi^H x.
\]

	extbf{Reproducibility hooks.} The QuantoniumOS repository includes:
\begin{itemize}
    \item Unitarity and reconstruction validation: \texttt{tests/validation/test\_phi\_frame\_normalization.py}
    \item Finite-$N$ benchmark: \texttt{benchmarks/rft\_phi\_frame\_benchmark.py}
    \item Large-$N$ ($N\le 4096$) coherence diagnostics: \texttt{benchmarks/rft\_phi\_frame\_asymptotics.py}
    \item Large-$N$ real-data coefficient statistics (ECG): \texttt{benchmarks/rft\_phi\_nudft\_realdata\_eval.py}
\end{itemize}

### 2025-12-15 papers/paper.tex
Update Canonical RFT definition to Gram-normalized basis and align docs/proofs

\begin{abstract}
Fast Fourier transforms (FFTs) and discrete cosine transforms (DCTs) sit at the core of today’s compression, filtering, and simulation stacks, but they lock us into integer harmonics on a flat circle and force a hard trade-off between sparsity, mixing, and reversibility. This work introduces the \emph{Resonance Fourier Transform} (RFT), a family of unitary, FFT-class transforms built from irrational (golden-ratio--scaled) phase progressions and braided, toroidal topologies. We formalize two complementary realizations: a canonical $\Phi$-RFT derived via QR factorization (theoretical reference) and a closed-form fast $\Phi$-RFT factored as $\Psi = D_\phi C_\sigma F$ with $O(N \log N)$ complexity. Ten ``irrevocable theorems'' numerically validate unitarity, twisted-convolution diagonalization, sparsity structure, and hybrid DCT+RFT behavior, with empirical spectral sparsity reaching $98.6\%$ at $N=512$ and maximum unitarity deviation below $10^{-14}$.

@@

The Resonance Fourier Transform (RFT) was introduced and formally analyzed in~\cite{minier2025rft} (DOI: 10.5281/zenodo.17712905) as a unitary family of transforms built on irrational, golden-ratio--based phase progressions and braided toroidal topologies, with a canonical QR-derived $\Phi$-RFT for theory and a closed-form fast $\Phi$-RFT for practice. This paper extends that foundational work by demonstrating RFT's practical application in hybrid compression, cryptographic mixing, and quantum-inspired computing. On top of this basis, we build: (i) a rigorously validated stack of $\Phi$-RFT theorems covering unitarity, complexity, sparsity, and twisted-convolution algebra; (ii) a hybrid DCT+RFT codec that beats single-basis baselines on mixed-structure signals; (iii) an RFT-SIS experimental lattice playground with measured avalanche/diffusion; and (iv) a unified C/FPGA execution path that removes Python from the hot loop and demonstrates that this resonance-based approach is not just mathematically consistent, but also system-level executable.

>>>>

\begin{abstract}
Fast Fourier transforms (FFTs) and discrete cosine transforms (DCTs) sit at the core of today’s compression, filtering, and simulation stacks, but they lock us into integer harmonics on a flat circle and force a hard trade-off between sparsity, mixing, and reversibility. This work introduces the \emph{Resonance Fourier Transform} (RFT), a family of unitary, FFT-class transforms built from irrational (golden-ratio--scaled) phase progressions and braided, toroidal topologies. We formalize two complementary realizations: a canonical $\Phi$-RFT derived via Gram-matrix normalization (theoretical reference) and a closed-form fast $\Phi$-RFT factored as $\Psi = D_\phi C_\sigma F$ with $O(N \log N)$ complexity. Ten ``irrevocable theorems'' numerically validate unitarity, twisted-convolution diagonalization, sparsity structure, and hybrid DCT+RFT behavior, with empirical spectral sparsity reaching $98.6\%$ at $N=512$ and maximum unitarity deviation below $10^{-14}$.

@@

The Resonance Fourier Transform (RFT) was introduced and formally analyzed in~\cite{minier2025rft} (DOI: 10.5281/zenodo.17712905) as a unitary family of transforms built on irrational, golden-ratio--based phase progressions and braided toroidal topologies, with a canonical Gram-normalized $\Phi$-RFT for theory and a closed-form fast $\Phi$-RFT for practice. This paper extends that foundational work by demonstrating RFT's practical application in hybrid compression, cryptographic mixing, and quantum-inspired computing. On top of this basis, we build: (i) a rigorously validated stack of $\Phi$-RFT theorems covering unitarity, complexity, sparsity, and twisted-convolution algebra; (ii) a hybrid DCT+RFT codec that beats single-basis baselines on mixed-structure signals; (iii) an RFT-SIS experimental lattice playground with measured avalanche/diffusion; and (iv) a unified C/FPGA execution path that removes Python from the hot loop and demonstrates that this resonance-based approach is not just mathematically consistent, but also system-level executable.
